---
title: "STAT 2060 HW 1"
author: "Benjamin Buss"
date: "1/22/2020"
output:
  word_document: default
  html_document: default
---
#Homework 1 - STAT 2060 - Benjamin Buss
1. (6 pts) Enter the following data in R and call it P1:
23,45,67,46,57,23,83,59,12,64

```{r}
P1 <- c(23,45,67,46,57,23,83,59,12,64)
```

a. What is the maximum value?
```{r}
max(P1)
```

b. What is the minimum value?
```{r}
min(P1)
```

c. What is the mean value?
```{r}
mean(P1)
```

2. (6 pts) Oh no! The next to last (9th) value was mistyped - it should be 42.
a. Change it, and see how the mean has changed.
```{r}
P2 <- c(23,45,67,46,57,23,83,59,42,64)
mean(P2)
mean(P2) - mean(P1)
```

b. How many values are greater than 40 in the correct data set?
```{r}
sum(P2 > 40)
```


c. What is the mean of values which are over 40 in the correct dataset?
```{r}
P3 <- c(45, 67, 46, 57, 83, 59, 42,64)
mean(P3)
```


3. (8 pts) Using the data from problem 2 find
a. the sum of P1
```{r}
sum(P2)
```

b. the mean (using the sum and length(P1))
```{r}
sum(P2)/length(P2)
```

c. the log(base10) of P1
```{r}
log10(P2)
```

d. the deviaions: the difference between each element of P1 and the mean of P1.
```{r}
mean(P2) - P2
```

