---
title: "STAT 2060 - BENJAMIN BUSS - HW 2(Portion 2)"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---

(5 pts) The Cornell Lab of Ornithology holds an annual Christmas Bird Count, in which bird watchers at various locations around the country see how many different species of birds they can spot. Here are some of the counts reported from sites in Texas during the 2010 event:
150  216  177  150  166  156  159  160  164  169  175  150  178  183  199  154  164  203  158  231

a. Save the data as csv file, and then use R function read.csv to read csv in to R. (1 pts)
```{r}
birds <- c(150, 216, 177, 150, 166, 156, 159, 160, 164, 169, 175, 150, 178, 183, 199, 154, 164, 203, 158, 231)
write.csv(birds, "birds.csv")
bird <- read.csv("~/class_work/birds.csv", row.names = 1)
```

b. Draw a histogram and give the title: The Species of Birds, and the subtitle: From the Cornell Lab of Ornithology (2 pts)
```{r}
hist(birds, main = "The Species of Birds", sub = "From the Cornell Lab of Ornithology")
```

c. Create a stem-and-leaf display of these data. (2 pts)
```{r}
stem(birds)
```

(6 pts) The sample of high temperatures for nineteen cities in the southeast (in degrees Fahrenheit) is in the following:
85  82  93  88  92  45  79  84  90  77  83  91  41  101  89  90  85  87  141
```{r}
temp <- c(85, 82, 93, 88, 92, 45, 79, 84, 90, 77, 83, 91, 41, 101, 89, 90, 85, 87, 141)
```

a. Find five-number summary and the inter-quartile range. (2pts)
```{r}
summary(temp)
IQR(temp)
```

b. Check potential outliers.(2pts)
```{r}
which(temp > 1.5 * summary(temp)[["1st Qu."]] & temp < 1.5 * summary(temp)[["3rd Qu."]])
```

c. Draw a boxplot and add the sample mean to the boxplot (2pts)
```{r}
boxplot(temp)
points(mean(temp), col = "red")
```

