---
title: "STAT 2060 - HW4(Part 1) - BENJAMIN BUSS"
output: html_notebook
---

(6 pts) The amount of time Americans spend using iphone is closely monitored by firms such as A.C. Nielsen because this helps to determine advertising pricing for commercials

a. (2 pts) According to American Time Use Survey, adult Americans spend 2.35 hours per day using iphone on a weekday. Assume that the standard deviation for “time spending using iphone on a weekday” is 1.93 hours. If a random sample of 40 adult Americans is obtained, describe the sampling distribution of  xbar, the mean amount of time spending iphone on a weekday. Use R to draw the density curve of xbar.
```{r}
data <- rnorm(40, 2.35, 1.93)

```


b. (2 pts) Determine the probability that a random sample of 40 adult Americans results in a mean time using iphone on a weekday of between 2 and 3 hours.
```{r}

```

c. (2 pts) Find the mean time of watching television with 98th percentile.
```{r}

```

2. (4 pts) A wildlife biologist examines frogs for a genetic trait he suspects may be linked to sensitivity to industrial toxins I the environment. Previous research had established that this trait is usually found in 1 of every 8 frogs. He collects and examines two-dozen frogs. If the frequency of the trait has not changed, what’s the probability he finds the trait in

a. At least 5 frogs?
```{r}

```

b.Ten or eleven frogs?
```{r}

```

