---
title: "STAT 2060 - HW3 - BENJAMIN BUSS"
output:
  word_document: default
  html_notebook: default
---

A researcher wants to know if the gestation period of an animal can be used to predict life expectancy. She collects the following data.  Use R to answer the following questions:

(2 pts) Save the following data as csv (or txt) file and import it into R.
```{r}
data <- read.csv("stat2060hw3data.csv")
```

(2 pts) Include the output of names(), str(), and head().
```{r}
names(data)
str(data)
head(data)
```

(2 pts) Draw a scatter plot.
```{r}
plot(data$Life.Expectancy, data$Gestation.Period)
```

(4 pts) Use both formula and R function to compute the linear correlation coefficient between gestation period and life expectancy.
```{r}
model <- lm(data$Gestation.Period ~ data$Life.Expectancy)
summary(model)

# Getting Linear Algebra Version
M = matrix(c(rep(1, 10),data$Life.Expectancy), byrow=F, ncol = 2)
N = matrix(data$Gestation.Period, byrow = F, ncol = 1)
B = solve(t(M) %*% M) %*% (t(M) %*% N)
B
```

(3 pts) Use both formula and R function to find the least-squares regression line to fit the sample data.
```{r}
# Gestation = -128.61197 + 20.200* Life.Expectancy
```

(2 pts) Draw a least squares regression line in the scatter plot.
```{r}
plot(data$Life.Expectancy, data$Gestation.Period) + abline(coefficients(model))
```

