---
title: "STAT 2060 - QUIZ 5 - BENJAMIN BUSS"
output:
  word_document: default
  html_notebook: default
---

Suppose that a student’s verbal score X from next year’s Graduate Record Exam can be considered an observation from a normal population having mean 547 and standard deviation 120.

(a)  (3 pts) Draw the density curve of X using the following function in R: seq(), dnorm(), and plot().
```{r}
#example <- dnorm(x = 1, mean = 547, sd = 120, n = 100)
x <- rnorm(100, 547, 120)
hist(x)

xfit = seq(min(x), max(x), length = 100)
yfit = dnorm(xfit, mean = mean(x), sd = sd(x))
#lines(xfit, yfit)
```

(b) (1 pts) Use a function in R to compute the probability that a randomly selected student has verbal score more than 600.
```{r}
z = (600 - 547)/(120)
pnorm(z, lower.tail = FALSE)
```

(c) (1 pts) Use a function in R to compute the Probability that a randomly selected student who's  score is below 400.
```{r}
z = (400 - 547)/(120)
pnorm(z, lower.tail = TRUE)
```


