---
title: "STAT 4100 HW 1 - Benjamin Buss"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---
```{r}
par(mfrow=c(2,2))
initial <- c(38, 58, 65, 74, 56, 36, 56, 52)
later   <- c(52, 60, 72, 72, 54, 48, 58, 60)
```
1. (5 pts) Read syllabus 

Done

2. (15 pts) Read 1.1.4 Randomization from chapter 1. Explain the concepts of experimenter bias and systematic bias in two examples that are different from the textbook. 

An example of experimenter bias could be when designing an experiment concerning the affect of a specific tool on a workers productivity. The experimenter unconsiously selects the best workers to test the tool, using the worst workers as the control.

Systematic bias may be present in an experiment where the experimenter is attempting to determine the temperature difference between two points using two different thermometers, one of the thermometers may be miscalibrated causing systematic error.


3. (80 pts) A Secchi disk is an 8-inch diameter weighted disk that is painted black and white and attached to a rope. The disk is lowered into water and the depth at which it is no longer visible is recorded. The measurement is an induction of water clarity. An environmental biologist is interested in determining whether the water clarity of the lake at Joliet Junior College is improving. She takes measurements at the same location on the same dates during the course of a year and repeats the measurements on the same dates 5 years later. She obtains the following results:
DataInitial depthDepth 5 years later1385225860365724747255654636487565885260

(a) (Total 30 pts) Identify the type of experimental design (5 pts); identify the response variable (5 pts); treatments (5 pts); describe the basic principles of experimental design applied to this design(15 pts).

Experiment design: replication       

Response variable: perceived depth           

Treatments: time between measurements 

Basic principlies: We are attempting to measure changes in visibility by checking the depth of the Secchi disk(Experimental Units), This is being affected by time(treatments). We are attempting to control for factors by replicating the experiment 8 times at the same location.


(b) (Total 10 pts) Provide the appropriate boxplot(s) and normality plot(s) based on the needs of the data analysis. 
```{r}
boxplot(initial, later)
qqnorm(initial)
qqnorm(later)
```

(c) (5 pts) Define the appropriate parameter, the mean difference, in words precisely for measuring the change of the initial depth to 5 years later depth. 

The mean difference is the difference in depth between the replicated values across the 5 year time span.
```{r}
mu <- later - initial
```


(d) (Total 15 pts) Give an estimate of the mean difference defined in part (c) and provide a 95% confidence interval for the mean difference (8 pts). What assumptions are needed for this confidence interval to be valid (5 pts)? Are the assumptions satisfied based on the results from part (b)? 
```{r}

xbar = mean(initial)
ybar = mean(later)
xvar = var(initial)
yvar = var(later)
# Find point estimate
pt = ybar -xbar

# Find a confidence interval
sp = sqrt((8*xvar + 8*yvar)/14)
lb = pt - qt(0.975, 14) * sp * sqrt(1/8 + 1/8) #Lower bound
ub = pt + qt(0.975, 14) * sp * sqrt(1/8 + 1/8) #Upper bound

mean(mu) # Mean of differences

print("A confidence interval would be +- 12.49")
```


(e) (Total 15 pts) State appropriate null and alternative hypotheses for examining the question of whether or not the course is associated with an improvement in clarity of the lake (5 pts). Perform a t-test to test this hypothesis assuming an α-level of 0.05. (10 pts)

null hypothesis        Mu1 - Mu2 = 0    # No change in visibility

alternative hypothesis Mu1 - Mu2 > 0    # Improvement to visibility

```{r}
t.test(later, initial, alternative = "less", con.level = 0.05, paired = T)
```


(f) (5 pts) What would be a causal interpretation of the conclusion in part (d)? Are you comfortable with this interpretation? Explain in a sentence or two.

There was an improvement in visibility over the 5 years measured.

